### 2023-06-12 draft_work.Rmd
Added "Draft" Notebook

---
title: "R Notebook"
output: html_notebook
---

```{r}
international_visits %>%
  group_by(year, quarter, country) %>%
  summarise(total_visits = sum(visits),
            total_spend = sum(spend)) %>% 
  ggplot(aes(x = total_visits, y = total_spend)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
demographics_all %>% 
  ggplot(aes(x = visits, y = expenditure)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}

ips_2002_2019 %>% 
  ggplot(aes(x = visits, y = spend)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Visits (Million)",
    y = "Expenditure (£Million) \n",
    title = "Scotland: International Visits vs. Spend",
    subtitle = "2012-2019")

```

```{r}

model <- lm(formula = spend ~ visits, data = ips_2002_2019)

model2 <- lm(formula = spend ~ visits + nights, data = ips_2002_2019)

```

```{r}
summary(model)
```

```{r}
summary(model2)
```

```{r}
library(modelr)
```

```{r}

ips_2002_2019 %>%
  add_predictions(model) %>%
  add_residuals(model) %>% 
  ggplot(aes(x = visits, y = spend)) +
  geom_point() +
  geom_line(aes(y = pred), col = "red") +
  annotate("text", x = 1675, y = 2500, label = "spend ~ visits") +
  annotate("text", x = 1930, y = 2300, label = "b1 = 0.81800 / p-value = 4.78e-06") +
  annotate("text", x = 2170, y = 2100, label = "R-Squared = 0.8876 / Residual Standard Error: 183.3")

```

```{r}
library(ggfortify)
```

```{r}
autoplot(model)
```