### 2023-06-13 draft_work.Rmd
Re-plotted Demographic Plots Based On %, Added International Models

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(broom)
library(factoextra)
```

```{r}
activities %>%
  group_by(tourism_activity) %>%
  mutate(exp_per_visit = Expenditure / Visits) %>% 
  summarise(mean_exp_per_visit = mean(exp_per_visit)) %>% 
  ggplot() +
  geom_col(aes(x = reorder(tourism_activity, mean_exp_per_visit), y = mean_exp_per_visit,
               fill = tourism_activity == "Shopping for items that you do not regularly buy"),
           show.legend = FALSE) +
  scale_fill_manual(values = c("FALSE" = "grey40",
                               "TRUE" = "forestgreen")) +
  labs(
    x = "\n Activity",
    y = "Mean Expenditure \n",
    title = "Mean Expenditure per Activity  2013-2019") +
  coord_flip()
  # theme(axis.text.x = element_text(angle = 45, hjust=1))
```

```{r}
activities %>% 
  mutate(is_2019 = if_else(date_code == 2019, "2019", "2013-2018")) %>% 
  group_by(tourism_activity, is_2019) %>%
  arrange(tourism_activity) %>% 
  mutate(exp_per_visit = Expenditure / Visits) %>% 
  summarise(mean_exp_per_visit = mean(exp_per_visit)) %>% 
  mutate(difference_exp_per_visit = mean_exp_per_visit - lag(mean_exp_per_visit),
         difference_exp_per_visit_pct = (mean_exp_per_visit - lag(mean_exp_per_visit)) / lag(mean_exp_per_visit) * 100,
         difference_exp_per_visit_label = if_else(difference_exp_per_visit > 0, "Increased", "Decreased")) %>% 
  filter(difference_exp_per_visit != is.na(difference_exp_per_visit)) %>% 
  ggplot(aes(x = reorder(tourism_activity, difference_exp_per_visit), y = difference_exp_per_visit)) +
  geom_col(aes(fill = difference_exp_per_visit_label), show.legend = FALSE) +
  # geom_text(aes(label = str_c(floor(difference_exp_per_visit_pct), "%"), vjust = -0.5)) +
  scale_fill_manual(values = c("Increased" = "forestgreen",
                               "Decreased" = "firebrick")) +
  labs(
    x = "\n Activity",
    y = "Mean Expenditure per Visit \n",
    title = "Mean Expenditure per Visit, 2019 vs. 2013-2018 Avg.") +
  coord_flip()
```

-----------

```{r}
activities_summary <- activities %>%
  group_by(tourism_activity) %>% 
  summarise(mean_expenditure = mean(Expenditure),
            mean_visits = mean(Visits),
            exp_per_visit = mean_expenditure / mean_visits) %>% 
  arrange(desc(exp_per_visit))
```

```{r}
activities_clust <- activities %>%
  group_by(tourism_activity) %>% 
  summarise(total_expenditure = sum(Expenditure),
            total_visits = sum(Visits)) %>% 
  select(total_expenditure, total_visits)
```

```{r}
min_k <- 1
max_k <- 5

k_clusters <- tibble(k = min_k:max_k) %>% 
  mutate(kclust = map(k, ~ kmeans(activities_clust, # Create K-means cluster beteen 1-25 clusters.
                                  centers = .x,
                                  nstart = 25)),
         tidied = map(kclust, tidy), # Tidies `kclust`
         glanced = map(kclust, glance), # Uses tidied to create a glance
         augmented = map(kclust, augment, activities_summary)) # Adds labels

# Unnest `glanced` to view `tot.withinss`
clusterings <- k_clusters %>% 
  unnest(glanced)

clusterings
```

```{r}
clusterings %>% 
  ggplot(aes(x = k, y = tot.withinss)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 25, 1))
```

```{r}
fviz_nbclust(activities_clust,
             kmeans,
             method = "wss", #within sum of squares
             nstart = 25)
```

```{r}
fviz_nbclust(activities_clust,
             kmeans,
             method = "silhouette",
             nstart = 25)
```

```{r}
clusterings %>% 
  unnest(augmented) %>% # Unnest augmented created in step 5.1
  filter(k == 2) %>% # Set number of clusters
  ggplot(aes(x = total_visits, y = total_expenditure, colour = .cluster)) +
  geom_point()
```

```{r}
clusterings %>% 
  unnest(augmented) %>% # Unnest augmented created in step 5.1
  filter(k == 2) %>% 
  group_by(.cluster)
```

-----------------------------------------------------------